Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-04-05 manuscript/manuscript.Rmd
Aggregated plotting data are now stored instead of raw data due to large file size.
Implemented empirical goodness of fit test into the package.

```
We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic. Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure *Insert*. Since for $\kappa \leq 0.80$ error curves are similar, we fix the value of the central portion $\kappa$ to 0.80. This leads to test statistic values listed in Table *insert*.

@@

.plot_mean_absolute_residual_error(
  plot_theme = plot_theme,
  manuscript_dir = manuscript_dir)

```
```{r include = FALSE, eval = FALSE}
data <- .get_test_statistic_data(manuscript_dir = manuscript_dir)

@@

@@

>>>>

```
We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic. Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure *Insert*. Since for $\kappa \leq 0.80$ error curves are similar, we fix the value of the central portion $\kappa$ to 0.80. This leads to test statistic values listed in Table *insert*, which derive from the type I error rates of the Box-Cox transformation.

@@

.plot_type_1_error_rate(
  plot_theme = plot_theme,
  manuscript_dir = manuscript_dir)

@@

```{r include = FALSE, eval = FALSE}
table_data <- .get_test_statistics_data(manuscript_dir = manuscript_dir)
```

```{r include = FALSE, eval = FALSE}
load(file = "./R/sysdata.rda")

gof_lookup_table <- .get_test_statistics_data(manuscript_dir = manuscript_dir, reduce = TRUE, as_table = FALSE)
gof_lookup_table[, "method" := NULL]

usethis::use_data(
  gof_lookup_table, 
  two_sided_function_parameters, 
  internal = TRUE, 
  overwrite = TRUE)
```

```{=latex}
\begin{table}
\begin{center}
\caption{Probability of central normality, based on type I error rates for Box-Cox transformations.}
\begin{tabular}{c c c c c c c}

\toprule
0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\

\midrule
0.029 & 0.039 & 0.047 & 0.054 & 0.062 & 0.069 & 0.090 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

@@

# Appendix D: Generalisability of empirical goodness of fit test

The empirical goodness of fit test was characterised using the empirical distribution of type I errors related to the mean residual error in the central portion ($\kappa = 0.80$) of the transformed feature distribution, in the presence of outliers. The corresponding power transformations were both robust and shift sensitive, and used maximum likelihood estimation (MLE) for optimisation of transformation parameters. An important concern is how well the test performs for conventional power transformations (in absence of outliers), and for optimisation criteria that are objectively better than MLE, i.e. the Cramér-von Mises criterion.

For Figure *insert* determined type 1 error rate as before, but for several variants. In addition to robust, shift sensitive transformations using MLE, conventional transformations and robust, shift-sensitive using the Crámer-von Mises criterion were assessed. Conventional transformations only used randomly drawn asymmetric generalised normal distributions without added outliers. This figure suggests two things: If conventional transformations were used instead of robust, shift sensitive variants, the test characteristics would be highly similar. However, the robust, shift sensitive variant shows that the test characteristics are even stable in the presence of outliers. Secondly, the optimisation criterion does directly matter, and compared to MLE, the Crámer-von Mises criterion results in transformed distributions that better approximate central normality. In this case, one could argue whether characteristics of the empirical goodness of fit test should be based on the Crámer-von Mises criterion instead of MLE. We would argue that MLE-based test characteristics form a good foundation, for two reasons. One, MLE is the most prevalent optimisation criterion for power transformations, and two, the resulting transformations still approach central normality.

```{r gof_test_type_1_error_rate_appendix, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Type I error rates for power transformation variants. ",
  "The robust, shift sensitive power transformation using maximum likelihood estimation (MLE) is shown as comparison. ",
  "Optimisation of transformation parameters using the Crámer-von Mises (C-vM) criterion leads to lower residual errors compared to MLE, ",
  "particularly for Box-Cox transformations, i.e. they lead to transformations that are closer to central normality. ",
  "Conventional transformations (in the absence of outliers) yield curves that are comparable to their robust, shift sensitive variants."
)

.plot_type_1_error_rate_appendix(
  plot_theme = plot_theme,
  manuscript_dir = manuscript_dir)
